---
# title: "2021-INFORMS"
# subtitle: "asd"
# author: "Nicolò Bertani"
# date: "10/18/2021"
output: 
  ioslides_presentation:
    incremental: false
    widescreen: true
    # smaller: true
    transition: 0
    css: style.css
    self_contained: true
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r includes}
setwd("~/OneDrive - ucp.pt/Shared Working Directory - ED, NB/ELICIT EXPERIMENT/R code and deliverables")
source("../Helper functions/M-spline and I-spline.R")
source("../Helper functions/Helper functions for data analysis.R")

cols <- c(
  rgb(.8,0,0),
  rgb(26/255,48/255,132/255),
  rgb(0,.6,0))

cols2 <- c(
  'steelblue',
  'orange',
  'springgreen4',
  'darkred',
  'dodgerblue'
)

cols3 <- c(
  'skyblue',
  'hotpink',
  'chartreuse2'
)
```


## Overview {.build .overview}

As experimenters, we often desire **faster** and **simpler** elicitation procedures that remain as **reliable** as longer alternatives.

We devise a fast and simple procedure, using simple geometric intuition to shorten questioning.

We suggest a functional form that is as tractable as the Neo-Additive, while improving descriptiveness over previous forms.

We validate our procedure by eliciting probability weighting in the lab.




## Central idea {.build}

Suppose I want to elicit a decision-theoretic function (utility, probability weighting, ...).

Can we iteratively incorporate preferences provided by a respondent to update our information and ask optimal questions? 

Some procedures take a Bayesian approach and update parameter distributions (e.g. Cavagnaro et alii 2013).

Differently, we:

- reduce the space where the itself function can lie, by defining and updating bounds on it.
- formulate this problem as an optimization problem.

In particular, here we consider Yaari's model (1987) and elicit the probability weighting function.




## Elicitation as an optimization problem {.build}

At every point in the support, we want to find the minimal or maximal value the function can take.

Thus, at every point, the upper bound can be defined as the solution to:
$$
\begin{align}
& \underset{}{\max} & & \text{function} 
\\
& \text{subject to} & & \text{functional assumptions (e.g. monotonicity)}
\\
& & & \text{previously collected preferences}
\end{align}
$$

<br>
The optimal question bisects the maximal distance between the upper and lower bound on the function.

The procedure iteratively updates the bounds and asks the optimal question.



## Functional form {.smaller .build}

For almost all standard functional forms, the optimization problem underlying the bounds cannot be solved.

We define the function as a convex combination of Integrated Spline bases. 

```{r splines, fig.align='center', fig.width=9, fig.height=3.6}
helper <- seq(0, 1, .01)
chosen.xi <- c(.1, .9)
order <- 3
m <- length(M_knots_sequence(order, chosen.xi)) - order
examples_lambda <- list(c(.2, .2, .2, .2, .2), 
                     c(.4, .3, .2, .1, 0),
                     c(0, .1, .2, .3, .4))
spline_bases <- I_spline(helper, order, interior_knots = chosen.xi, individual = T)
spline_examples <- sapply(examples_lambda, function(l) I_spline(helper, order, interior_knots = chosen.xi, individual = F, lambdas = l))

par(
  mfrow = c(1, 2),
  xpd = T,
  mar = c(3.2, 3, 1.6, 0),
  oma = c(0, 0, 0, 0) + .1,
  xaxs = "i",
  yaxs = "i",
  pty = "s"
)

s()
for (i in seq(ncol(spline_bases))) {
  lines(spline_bases[, i] ~ helper, col = cols2[i], lw = 3)
}
mtext(expression(p), 1, 2.4, cex = 1.2)
mtext(expression(w(p)), 2, 2, cex = 1.2)
mtext('Ispline bases', 3, .6, cex = 1.2, font = 2)

s()
for (i in seq(ncol(spline_examples))) {
  lines(spline_examples[, i] ~ helper, col = cols3[i], lw = 3)
}
mtext(expression(p), 1, 2.4, cex = 1.2)
mtext(expression(w(p)), 2, 2, cex = 1.2)
mtext('Examples of combinations', 3, .6, cex = 1.2, font = 2)
# 
```

Based on this functional form, the bounds can be found using solvable linear programs (LPs).





## The procedure at work

```{r}
shinyAppFile("elicit/app.R")
```



## Procedure performance {.build}

Updates take about half a second to compute.

Experimentally, a function is elicited with 10 binary choices on average.

Experiments last on average 6.5 minutes, including instructions.



## Thank you

&nbsp;
<div style="font-weight:bold; font-size: xx-large; text-align:center;">
</div>
&nbsp;
<div style="font-weight:bold; font-size: xx-large; text-align:center;">
</div>
&nbsp;
<div style="font-weight:bold; font-size: xx-large; text-align:center;">
</div>
&nbsp;
<div style="font-weight:bold; font-size: xx-large; text-align:center;">
</div>
&nbsp;
<div style="font-weight:bold; font-size: xx-large; text-align:center;">
</div>
&nbsp;
<div style="font-weight:bold; font-size: xx-large; text-align:center;">
</div>
&nbsp;
<div style="text-align: center; font-size: 18px">
Manuscript available on SSRN:

[https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3569625](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3569625)
</div>
&nbsp;
<div style="text-align: center; font-size: 18px">
Presentation available on GitHub:

[https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3569625](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3569625)
</div>
